---
title: "Prediction Assignment"
output: html_document
---
Gitta Borghart
28-04-2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The project can be found on github: 

## Source of Data 
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

```{r}
setwd("~/")

# Load the training dataset
Training <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=T)

# Load the testing dataset
Testing <- read.csv("pml-testing.csv", na.strings=c("NA",""), strip.white=T)

#Data cleaning
features <- names(Testing[,colSums(is.na(Testing)) == 0])[8:59]

# Only use features used in testing cases.
Training <- Training[,c(features,"classe")]
Testing <- Testing[,c(features,"problem_id")]

dim(Training)
dim(Testing)


library(caret)
set.seed(12345)
inTrain <- createDataPartition(Training$classe, p=0.60, list=FALSE)
train1  <- Training[inTrain,]
test1  <- Training[-inTrain,]

dim(train1)
dim(test1)


#The model
library(rpart)
library(rpart.plot)
library(rattle)
modeltree <- rpart(classe ~ ., data = train1, method="class")
fancyRpartPlot(modeltree)


#Predicting
set.seed(12345)

prediction <- predict(modeltree, test1, type = "class")
confusionMatrix(prediction, test1$class)


#Check for the Out of Sample Error calculation
Check <- function(values, predicted) {
  sum(predicted != values) / length(values)
}
OutofSample_ErrorRate <- Check(test1$classe, prediction)
print(OutofSample_ErrorRate)

```

Conclusion: Accuracy of prediction with the above data set is around 75%, which is quite alright. The out of sample error is 25%. 